---
title: "One Variable: Quantitative"
author: Orville D. Hombrebueno
date: today
format: html
editor: visual
toc: true
---

# The Data

```{r}
#| label: preliminaries
#| warning: false
#| message: false

# load packages
library(tidyverse)
library(knitr)
library(gtsummary)

# load the data
data(iris)
```

## Edgar Anderson's Iris Data

This famous (Fisher's or Anderson's) iris data set gives the measurements in centimeters of the variables sepal length and width and petal length and width, respectively, for 50 flowers from each of 3 species of iris. The species are Iris setosa, versicolor, and virginica.

The first 6 rows of the iris dataset is shown in @tbl-iris. To see more details about the data set, type `?iris` in the console.

```{r}
#| label: tbl-iris
#| warning: false
#| message: false
#| tbl-cap: "The first 6 observations in the iris dataset."

iris |> # loading the data
  head() |> # getting the first six rows of the data
  kable( # creating a table
    col.names = c( # changing the column names
      "Sepal Length",
      "Sepal Width",
      "Petal Length",
      "Petal Width",
      "Species"
    )
  )
```

## Descriptive Statistics

The mean, standard deviation, minimum, and maximum of the dataset by species is shown in @tbl-summary.

```{r}
#| label: tbl-summary
#| tbl-cap: "Summary statistics for the iris dataset."
#| message: false

iris |> 
  tbl_summary(
    include = c(Sepal.Length, Sepal.Width, Petal.Length, Petal.Width, Species),
    statistic = all_continuous() ~ "{mean} ± {sd} ({min} - {max})",
    digits = all_continuous() ~ 2,
    label = list(
      Sepal.Length ~ "Sepal Length",
      Sepal.Width ~ "Sepal Width",
      Petal.Length ~ "Petal Length",
      Petal.Width ~ "Petal Width"
    )
  )
```

# Hypothesis Testing

## Research Question

Is the sepal length 

## Hypothesis and Test Statistics

## Results and Interpretation

Hi
