---
title: "Two Variables: Qualitative"
author: Orville D. Hombrebueno
date: today
format: 
  docx:
    reference-doc: excel-report-template.docx
editor: visual
toc: true
cache: true
execute: 
  echo: false
---

# The Data

```{r}
#| label: preliminaries
#| warning: false
#| message: false

# load packages
library(tidyverse) 
library(knitr)
library(gtsummary)
library(ggstatsplot)
library(downlit)
library(broom)

# load the data
titanic <- ggstatsplot::Titanic_full |>
  mutate(
    Class = factor(Class)
  )
```

## The Titanic Data Set

The titanic data set is a data frame with 2201 rows and 5 variables.

-   `id` Dummy identity number for each person.

-   `Class` 1st, 2nd, 3rd, Crew.

-   `Sex` Male, Female.

-   `Age` Child, Adult.

-   `Survived` No, Yes.

This data set provides information on the fate of passengers on the fatal maiden voyage of the ocean liner Titanic, summarized according to economic status (class), sex, age and survival.

The first 6 rows of the titanic data set are shown in @tbl-titanic.

```{r}
#| label: tbl-titanic
#| warning: false
#| message: false
#| tbl-cap: "The first 6 observations of the data set."

titanic |>
  head() |> # getting the first six rows of the data
  kable() # creating a table
```

## Descriptive Statistics

The frequency distribution of `Survived` when grouped according to `Sex` is shown in @tbl-summary. Most of the passengers are male. The majority of female passengers survived, whereas the majority of male passengers did not.

```{r}
#| label: tbl-summary
#| tbl-cap: "Summary statistics of `Survived` when grouped according to `Sex`."
#| message: false

titanic |> 
  tbl_summary(
    by = Sex,
    include = Survived
  )
```

# Hypothesis Testing

## Research Question

Does a passenger's sex affect their survival status?

## Hypothesis and Test Statistics

-   $H_0:p_{Survived}-p_{Sex}=0$; passenger's sex does not affect their survival status.

-   $H_a:p_{Survived}-p_{Sex}\ne0$; passenger's sex affects their survival status.

-   Significance level, $\alpha=5\%$.

-   **Pearson's $\chi^2$-test of independence**

-   Independence.

-   Expected counts: all expected counts are $\ge5$ (see @tbl-expected).

```{r}
#| label: tbl-expected
#| tbl-cap: Expected counts.

# Create contingency table
observed <- table(titanic$Survived, titanic$Sex)

# Perform chi-squared test
chi_test <- chisq.test(observed)

# Extract the expected counts
expected <- chi_test$expected

# View expected counts
expected |>
  as_tibble() |>
  mutate(
    " " = c("No", "Yes")
  ) |>
  relocate(" ", .before = 1) |>
  mutate(across(where(is.numeric), ~ round(.x, 2))) |>
  kable()
```

## Results and Interpretation

The test statistics is shown on @tbl-stats.

```{r}
#| label: tbl-stats
#| tbl-cap: Test statistics.

chi_test |>
  tidy() |>
  select(statistic, p.value, parameter, method) |>
  rename(df = parameter) |>
  mutate(across(where(is.numeric), ~ round(.x, 2))) |>
  kable()
```

Based on the *p*-value being lesser than 5% (*p*-value = .001), we reject the null hypothesis. There is a strong evidence that passenger's sex affects their survival status.

