---
title: "One Variable: Qualitative"
author: Orville D. Hombrebueno
date: today
format: 
  docx:
    reference-doc: excel-report-template.docx
editor: visual
toc: true
cache: true
execute: 
  echo: false
---

# The Data

```{r}
#| label: preliminaries
#| warning: false
#| message: false

# load packages
library(tidyverse) 
library(knitr)
library(gtsummary)
library(ggstatsplot)
library(downlit)
library(broom)

# load the data
titanic <- ggstatsplot::Titanic_full |>
  mutate(
    Class = factor(Class)
  )
```

## The Titanic Data Set

The `Titanic_full` data set of the [{ggstatsplot}]{style="font-family:monospace;"} package is a data frame with 2201 rows and 5 variables.

-   id. Dummy identity number for each person.

-   Class. 1st, 2nd, 3rd, Crew.

-   Sex. Male, Female.

-   Age. Child, Adult.

-   Survived. No, Yes.

This data set provides information on the fate of passengers on the fatal maiden voyage of the ocean liner Titanic, summarized according to economic status (class), sex, age and survival.

The first 6 rows of the titanic data set are shown in @tbl-titanic.

```{r}
#| label: tbl-titanic
#| warning: false
#| message: false
#| tbl-cap: "The first 6 observations of the data set."

titanic |>
  head() |> # getting the first six rows of the data
  kable() # creating a table
```

## Descriptive Statistics

The frequency distribution of the variables in the data set is shown in @tbl-summary. Most of the passengers are crew members, male, and adult. Only 32% of the passengers survived.

```{r}
#| label: tbl-summary
#| tbl-cap: "Summary statistics of the data set."
#| message: false

titanic |> 
  tbl_summary(
    include = c(Class, Sex, Age, Survived)
  )
```

A visualization of the `Survived` variable is shown in @fig-viz.

```{r}
#| label: fig-viz
#| fig-cap: Bar chart
#| fig-height: 3

ggplot(titanic, aes(x=Survived)) + 
  geom_bar() +
  theme_minimal()
```

# Hypothesis Testing

::: panel-tabset \## Two-Sided

### Research Question

Do the passengers have a 50/50 chance of surviving?

### Hypothesis and Test Statistics

-   $H_0:p=50\%$; proportions of surviving are equal.

-   $H_a:p\ne50\%$; proportions of surviving are different.

-   Significance level, $\alpha=5\%$.

-   **one proportion z-test**

-   Success-failure condition is met.

### Results and Interpretation

The test statistics is shown on @tbl-stats-ts.

```{r}
#| label: tbl-stats-ts
#| tbl-cap: Test statistics.

prop.test(
  x = 711, # number of successes
  n = 2201, # total number of trials (711 + 1490)
  p = 0.35,
  alternative = "two.sided"
) |>
  tidy() |>
  select(estimate, statistic, p.value, method, alternative) |>
  mutate(across(where(is.numeric), ~ round(.x, 2))) |>
  kable()
```

Based on the *p*-value being less than 5% (*p*-value = .001), we reject the null hypothesis. There is a strong evidence that the proportion of surviving is different among passengers. This may be influenced by certain factors like class, gender, or age?

\pagebreak

## One-Sided

### Research Question

Is the survival rate less than the expected survival rate of 35%?

### Hypothesis and Test Statistics

-   $H_0:p=35\%$; the survival rate is equal to the expected survival rate of 35%.

-   $H_a:p<35\%$; the survival rate is less than the expected survival rate of 35%.

-   Significance level, $\alpha=5\%$.

-   **one proportion Z-test**

-   Success-failure condition is met.

### Results and Interpretation

The test statistics is shown in @tbl-stats-os.

```{r}
#| label: tbl-stats-os
#| tbl-cap: Test statistics.

prop.test(
  x = 711, # number of successes
  n = 2201, # total number of trials (711 + 1490)
  p = 0.35,
  alternative = "less"
) |>
  tidy() |>
  select(estimate, statistic, p.value, method, alternative) |>
  mutate(across(where(is.numeric), ~ round(.x, 2))) |>
  kable()
```

Based on the results and the *p*-value being lesser than 5% (*p*-value = 0.004), we reject the null hypothesis. There is strong evidence that the survival rate is less than the expected survival rate of 35%.
