---
title: "One Variable: Quantitative"
author: Orville D. Hombrebueno
date: today
format: 
  docx:
    reference-doc: excel-report-template.docx
editor: visual
toc: true
cache: true
execute: 
  echo: false
---

# The Data

```{r}
#| label: preliminaries
#| warning: false
#| message: false

# install packages: install.packages("package")
# load packages
library(tidyverse) 
library(knitr)
library(gtsummary)
library(ggstatsplot)
library(downlit)
library(broom)

# load data
data(iris)
```

## Edgar Anderson's Iris Data

This famous (Fisher's or Anderson's) [iris](https://onlinelibrary.wiley.com/doi/10.1111/j.1469-1809.1936.tb02137.x) data set gives the measurements in centimeters of the variables sepal length and width and petal length and width, respectively, for 50 flowers from each of 3 species of iris. The species are Iris setosa, versicolor, and virginica.

The first 6 rows of the iris data set are shown in @tbl-iris.

```{r}
#| label: tbl-iris
#| warning: false
#| message: false
#| error: false
#| tbl-cap: "The first 6 observations of the data set."

iris |> # call the data
  head() |> # get the first six rows of the data
  kable( # create a table
    col.names = c( # change the column names
      "Sepal Length",
      "Sepal Width",
      "Petal Length",
      "Petal Width",
      "Species"
    )
  )
```

## Descriptive Statistics

The mean, standard deviation, minimum, and maximum of the data set by species are shown in @tbl-summary. There is a total of 150 observations; 50 observations per species.

```{r}
#| label: tbl-summary
#| tbl-cap: "Summary statistics of the data set."
#| message: false

iris |> # call the data
  tbl_summary( # create a table
    include = c(Sepal.Length, Sepal.Width, Petal.Length, Petal.Width, Species), # select variables
    statistic = all_continuous() ~ "{mean} ± {sd} ({min} - {max})", # descriptive statistics
    digits = all_continuous() ~ 2, # no. of digits (rounding)
    label = list( # variable names
      Sepal.Length ~ "Sepal Length",
      Sepal.Width ~ "Sepal Width",
      Petal.Length ~ "Petal Length",
      Petal.Width ~ "Petal Width"
    )
  )
```

A visualization of sepal length is shown in @fig-viz.

```{r}
#| label: fig-viz
#| fig-cap: Histogram of sepal length.
#| fig-height: 3

iris |>
ggplot(aes(x = Sepal.Length)) +
  geom_histogram(
    binwidth = 0.5,
    fill = "blue",
    color = "white"
  ) +
  labs( 
    x = "Sepal Length", 
    y = "Frequency"
  ) +
  theme_minimal()

```

# Hypothesis Testing

::: panel-tabset
## Two-Sided

### Research Question

Is the mean sepal length different from the expected average of 5.8 cm?

### Hypothesis and Test Statistics

-   $H_0:\mu=5.8$; the mean sepal length is equal to 5.8 cm.

-   $H_a:\mu\ne5.8$; the mean sepal length is different from 5.8 cm.

-   Significance level, $\alpha=5\%$.

-   **one sample t-test**

-   The variance is unknown.

-   $n>30$; large sample.

### Results and Interpretation

The test statistics is shown on @tbl-stats-ts.

```{r}
#| label: tbl-stats-ts
#| tbl-cap: Test statistics.

t.test(
  iris$Sepal.Length, # selecting sepal length
  mu = 5.8, # hypothesized value of the mean in the population
  alternative = "two.sided" # other options: "two.sided" or "less"
) |>
  tidy() |>
  select(estimate, statistic, p.value, method, alternative) |>
  mutate(across(where(is.numeric), ~ round(.x, 2))) |>
  kable()
```

Based on the *p*-value being greater than 5% (*p*-value = 0.52), we do not reject the null hypothesis. There is no sufficient evidence that the mean sepal length is different from 5.8 cm.

\pagebreak

## One-Sided

### Research Question

Is the mean sepal length greater than the expected average of 5.7 cm?

### Hypothesis and Test Statistics

-   $H_0:\mu=5.7$; the mean sepal length is equal to 5.7 cm.

-   $H_a:\mu>5.7$; the mean sepal length is greater than 5.7 cm.

-   Significance level, $\alpha=5\%$.

-   **one sample t-test**

-   The variance is unknown.

-   $n>30$; large sample.

### Results and Interpretation

The test statistics is shown in @tbl-stats-os.

```{r}
#| label: tbl-stats-os
#| tbl-cap: Test statistics.

t.test(
  iris$Sepal.Length, # selecting sepal length
  mu = 5.7, # hypothesized value of the mean in the population
  alternative = "greater" # other options: "two.sided" or "less"
) |>
  tidy() |>
  select(estimate, statistic, p.value, method, alternative) |>
  mutate(across(where(is.numeric), ~ round(.x, 2))) |>
  kable()
```

Based on the results and the *p*-value being lesser than 5% (*p*-value = 0.02), we reject the null hypothesis. There is sufficient evidence that the mean sepal length is greater than the expected average of 5.7 cm.
:::
