---
title: "The Iris Dataset"
author: Orville D. Hombrebueno
date: today
format:
  html:
    embed-resources: true
    smooth-scroll: true
editor: visual
toc: true
code-link: true
code-fold: true
cache: true
---

# The Data

```{r}
#| label: preliminaries
#| warning: false
#| message: false

# install packages: install.packages("package")
# load packages
library(tidyverse) 
library(knitr)
library(gtsummary)
library(ggstatsplot)
library(downlit)

# load data
data(ChickWeight)
```

## Edgar Anderson's Iris Data

The `ChickWeight` data frame has 578 rows and 4 columns from an experiment on the effect of diet on early growth of chicks.

- weight. a numeric vector giving the body weight of the chick (gm).

- Time. a numeric vector giving the number of days since birth when the measurement was made.

- Chick. an ordered factor with levels 18 < ... < 48 giving a unique identifier for the chick. The ordering of the levels groups chicks on the same diet together and orders them according to their final weight (lightest to heaviest) within diet.

- Diet. a factor with levels 1, ..., 4 indicating which experimental diet the chick received.

The first 6 rows of the iris data set is shown in @tbl-iris.

```{r}
#| label: tbl-cw
#| warning: false
#| message: false
#| error: false
#| tbl-cap: "The first 6 observations of the data set."

ChickWeight |> # call the data
  head() |> # get the first six rows of the data
  kable( # create a table
    col.names = c( # change the column names
      "Weight",
      "Time",
      "Chick",
      "Diet"
    )
  )
```

## Descriptive Statistics

The mean, standard deviation, minimum, and maximum of the data set by species is shown in @tbl-summary. There are a total of 150 observations; 50 observations per species.

```{r}
#| label: tbl-summary
#| tbl-cap: "Summary statistics of the data set."
#| message: false

ChickWeight |> 
  tbl_summary()
```

# Hypothesis Testing

::: panel-tabset
## Two-Sided

### Research Question

Is the mean sepal length different from the expected average of 5.8 cm?

### Hypothesis and Test Statistics

-   $H_0:\mu=5.8$; the mean sepal length is equal to 5.8 cm.

-   $H_a:\mu\ne5.8$; the mean sepal length is different from 5.8 cm.

-   Significance level, $\alpha=5\%$.

-   **one sample t-test**

-   The variance is unknown.

-   $n>30$; large sample.

### Results and Interpretation

The histogram and test statistics is shown on @fig-test. Based on the histogram and the *p*-value being greater than 5% (*p*-value = 0.52), we do not reject the null hypothesis. There is no sufficient evidence that the mean sepal length is different from 5.8 cm.

```{r}
#| label: fig-test
#| fig-cap: Histogram and test statistics.

gghistostats( # create figure
  data = iris, # the data
  x = Sepal.Length, # variable of interest
  type = "parametric", # for student's t-test
  test.value = 5.8, # default value is 0
  xlab = "sepal length" # label of x-axis
) +
  labs(caption = NULL) # remove caption
```

## One-Sided

### Research Question

Is the mean sepal length greater than the expected average of 5.7 cm?

### Hypothesis and Test Statistics

-   $H_0:\mu=5.7$; the mean sepal length is equal to 5.7 cm.

-   $H_a:\mu>5.7$; the mean sepal length is greater than 5.7 cm.

-   Significance level, $\alpha=5\%$.

-   **one sample t-test**

-   The variance is unknown.

-   $n>30$; large sample.

### Results and Interpretation

```{r}
#| label: test-os

t.test(
  iris$Sepal.Length, # selecting sepal length
  mu = 5.7, # hypothesized value of the mean in the population
  alternative = "greater" # other options: "two.sided" or "less"
)
```

The test statistics is shown above. Based on the results and the *p*-value being lesser than 5% (*p*-value = 0.02), we reject the null hypothesis. There is sufficient evidence that the mean sepal length is greater than the expected average of 5.7 cm.
:::
