---
title: "One Variable: Quantitative"
author: Orville D. Hombrebueno
date: today
format:
  html:
    embed-resources: true
editor: visual
toc: true
code-link: true
code-fold: true
cache: true
---

# The Data

```{r}
#| label: preliminaries
#| warning: false
#| message: false

# load packages
library(tidyverse) 
library(knitr)
library(gtsummary)
library(ggstatsplot)
library(downlit)

# load the data
data(iris)
```

## Edgar Anderson's Iris Data

This famous (Fisher's or Anderson's) [iris](https://onlinelibrary.wiley.com/doi/10.1111/j.1469-1809.1936.tb02137.x) data set gives the measurements in centimeters of the variables sepal length and width and petal length and width, respectively, for 50 flowers from each of 3 species of iris. The species are Iris setosa, versicolor, and virginica.

The first 6 rows of the iris data set is shown in @tbl-iris.

```{r}
#| label: tbl-iris
#| warning: false
#| message: false
#| tbl-cap: "The first 6 observations in the iris dataset."

iris |> # loading the data
  head() |> # getting the first six rows of the data
  kable( # creating a table
    col.names = c( # changing the column names
      "Sepal Length",
      "Sepal Width",
      "Petal Length",
      "Petal Width",
      "Species"
    )
  )
```

## Descriptive Statistics

The mean, standard deviation, minimum, and maximum of the data set by species is shown in @tbl-summary. There are a total of 150 observations.

```{r}
#| label: tbl-summary
#| tbl-cap: "Summary statistics for the iris dataset."
#| message: false

iris |> 
  tbl_summary(
    include = c(Sepal.Length, Sepal.Width, Petal.Length, Petal.Width, Species),
    statistic = all_continuous() ~ "{mean} ± {sd} ({min} - {max})",
    digits = all_continuous() ~ 2,
    label = list(
      Sepal.Length ~ "Sepal Length",
      Sepal.Width ~ "Sepal Width",
      Petal.Length ~ "Petal Length",
      Petal.Width ~ "Petal Width"
    )
  )
```

# Hypothesis Testing

## Research Question

Is the mean sepal length in the dataset significantly different from the expected average of 5.8 cm?

## Hypothesis and Test Statistics

-   $H_0:\mu=5.8$

-   $H_1:\mu\ne5.8$

-   **one sample t-test**

-   The variance is unknown. 

## Results and Interpretation

The histogram and test statistics for sepal length is shown on @fig-test. The *p*-value is displayed. Based on this plot and the *p*-value being greater than 5% (*p*-value = 0.52), we do not reject the null hypothesis. There is no significant difference between the mean sepal length of the data set and the expected average of 5.8 cm.

```{r}
#| label: fig-test
#| fig-cap: Histogram and test statistics for sepal length.

# plot with stat test
gghistostats(
  data = iris, # dataframe from which variable is to be taken
  x = Sepal.Length, # numeric variable whose distribution is of interest
  type = "parametric", # for student's t-test
  test.value = 5.8, # default value is 0
  xlab = "sepal length"
) +
  labs(caption = NULL) # remove caption
```
